Runtime-Assisted Convergence in Replicated Data Types

### 2021-11-19 abstract-semantics.tex
Implementation section

\begin{definition}[Ancestor]
  In version history DAG $G = (V,E)$, version $v_0 \in V$ is said
  to be a (causal) ancestor of $v_1 \in V$ iff $v_0 \reaches v_1$.
  Versions $v_0, v_1 \in V$ are causally related iff either $v_0
  \reaches v_1$ or $v_1 \reaches v_0$.
\end{definition}

>>>>

\begin{definition}[Ancestor]
  \label{def:ancestor}
  In version history DAG $G = (V,E)$, version $v_0 \in V$ is said
  to be a (causal) ancestor of $v_1 \in V$ iff $v_0 \reaches v_1$.
  Versions $v_0, v_1 \in V$ are causally related iff either $v_0
  \reaches v_1$ or $v_1 \reaches v_0$.
\end{definition}

### 2021-11-19 concrete-semantics.tex
Edits to abstract and introduction
Implementation section

The rule \rulelabel{Sync} captures asynchronous communication between
a pair of replicas ($i$ and $j$) to get one of them ($i$) up-to-date
with the other ($j$).  Unlike the other rules, \rulelabel{Sync} does
not extend the version history graph, and therefore has no counterpart
in Fig.~\ref{fig:git-semantics}. It merely updates replica $i$'s
knowledge of branch $b_k$ if replica $j$ happens to have later updates
from $k$, i.e., $H_j(b_k)$ happens to be ahead of $H_i(b_k)$. The new
versions on $b_k$ known to $j$ but not $i$ are then simply transferred
to $i$.

>>>>

The rule \rulelabel{Sync} captures asynchronous communication between
a pair of replicas ($i$ and $j$) to get one of them ($i$) up-to-date
with the other ($j$).  Unlike the other rules, \rulelabel{Sync} does
not extend the version history graph, and therefore has no counterpart
in Fig.~\ref{fig:git-semantics}. It merely updates replica $i$'s
knowledge of branch $b_k$ if replica $j$ happens to have later updates
from $k$, i.e., $H_j(b_k)$ happens to be ahead of $H_i(b_k)$. The new
versions on $b_k$ known to $j$ but not $i$ are then simply replicated
at $i$.

\paragraph{Garbage Collection} One downside of the distributed
semantics in Fig.~\ref{fig:quark-semantics} is that maps $B$ and $N$
that together track the version history grow monotonically as the
execution progresses and new versions are created. Fortunately, this
is easy to address as the execution only needs version history
inasmuch as it is required to obtain the LCA versions for
\rulelabel{Merge}. Since LCA version vectors monotonically increase,
older versions with vectors less than the least known LCA vector can
simply be garbage collected. Moreover, a replica can make this
decision locally without having to synchronize with its peers. In
practice, applications may prefer to flush out older version history
to a stable storage from where it can be re-created as necessary.

### 2021-11-19 introduction.tex
Implementation section

% hinders the widespread adoption of replicated data
% types in software development.
In this paper we propose such an alternative approach to convergent
replicated data types. The key distinguishing characteristic of our
approach is the utilization of a distributed runtime to orchestrate
\emph{well-formed} distributed executions that are guaranteed to
converge. We define well-formedness as a constraint over the
\emph{states} that arise during the execution as opposed to the
\emph{operations} that are involved in the execution. To allow for its
enforcement, we adopt a state-centric model of replication based on
version-controlled \emph{mergeable} states~\cite{mrdt} instead of an
operation-centric model based on commutative operations. Unlike
commutativity, mergeability does not require a data type definition to
be refactored to suit distributed execution. Instead, the type
definitions needs to be \emph{extended} with a \C{merge} function to
reconcile concurrent versions of the type in presence of their common
ancestor version. Notably, \C{merge} function is left completely
unconstrained. Consequently, programmer can chose any merge semantics
that make sense within the context of the data type's (sequential)
semantics without risking the loss of convergence. Such strong
guarantees are made possible by the simplicity of state-based
reasoning combined with the run-time enforcement of well-formedness.

@@

  \item We formalize \quark distributed machine that implements the
    semantics of \quark abstract machine in the context of an
    asynchronous distributed system. The formalization addresses
    several practical concerns of distribution and serves as a
    blueprint for an efficient implementation.

>>>>

% hinders the widespread adoption of replicated data
% types in software development.
In this paper we propose such an alternative approach to convergent
replicated data types. The key distinguishing characteristic of our
approach is the utilization of a distributed runtime to orchestrate
\emph{well-formed} distributed executions that are guaranteed to
converge. We define well-formedness as a constraint over the
\emph{states} that arise during the execution as opposed to the
\emph{operations} that are involved in the execution. To allow for its
enforcement, we adopt a state-centric model of replication based on
version-controlled \emph{mergeable} states~\cite{mrdt} instead of an
operation-centric model based on commutative operations. Unlike
commutativity, mergeability does not require a data type definition to
be refactored to suit distributed execution. Instead, the type
definition only needs to be extended with a \C{merge} function to
reconcile concurrent versions of the type in presence of their common
ancestor version. Notably, \C{merge} function is left completely
unconstrained. Consequently, programmer can chose any merge semantics
that make sense within the context of the data type's (sequential)
semantics without risking the loss of convergence. Such strong
guarantees are made possible by the simplicity of state-based
reasoning combined with the run-time enforcement of well-formedness.

@@

  \item We formalize \quark distributed machine that implements the
    semantics of the \quark abstract machine in the context of an
    asynchronous distributed system. The formalization addresses
    several practical concerns of distribution and serves as a
    blueprint for an efficient implementation.

### 2021-11-19 motivation.tex
Implementation section

\noindent\paragraph{\emph{Add-Wins} Set CRDT} Unfortunately, \C{Set}
data type of Fig.~\ref{fig:cc-set} still admits divergent executions
due to concurrent updates.  Fig.~\ref{fig:crdt-execs-2} describes one
such execution. Here, replicas $R_1$ and $R_2$ both start with a
singleton set $\{e\}$.  Two distinct clients connect to $R_1$ and
$R_2$ respectively and issue two concurrent \C{remove($e$)}
operations. Later, another client connects to $R_2$ and issues an
\C{add($e$)} operation. The effects of these operations are
asynchronously applied at remote replicas as shown in the figure,
resulting in the divergent states at $R_1$ and $R_2$.

>>>>

\noindent\paragraph{Add-Wins Set CRDT} Unfortunately, \C{Set}
data type of Fig.~\ref{fig:cc-set} still admits divergent executions
due to concurrent updates.  Fig.~\ref{fig:crdt-execs-2} describes one
such execution. Here, replicas $R_1$ and $R_2$ both start with a
singleton set $\{e\}$.  Two distinct clients connect to $R_1$ and
$R_2$ respectively and issue two concurrent \C{remove($e$)}
operations. Later, another client connects to $R_2$ and issues an
\C{add($e$)} operation. The effects of these operations are
asynchronously applied at remote replicas as shown in the figure,
resulting in the divergent states at $R_1$ and $R_2$.
